Commutative Character Sheaves on smooth group schemes over finite fields

### 2016-01-13 followup.tex
Moved the lemma on geometrization of linear characters to its own section, at the beginning of the Applications.
Work on geometrization. Going home now.

@@

For use below, we recall some basic facts about the Heisenberg-Weil representation.

@@

We may now revisit the ingredients in the construction of the representation $\,^\circ \rho$ of $\underline{G}(\OK)$ along the lines indicated by Yu and recalled at the beginning of this section.
\begin{enumerate}
\item The continuous representation $\varrho$ of $\,^\circ K^0$ is a representation of $\underline{G}^0(\OK)$ obtained by inflation along $\underline{G}^0(\OK) \to \underline{G}^0(\Fq)$. 
Let $\sigma$ be that representation of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\sigma$ is itself obtained by pulling back a representation $\sigma^0$ along the $\Fq$-points of the quotient $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
\item Quasicharacters $\varphi_i : \underline{G}^i(\OK) \to \CC^\times$, for $i=0, \ldots d$.
\item The representation $h_{i+1}^*(W_{i+1})$ of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK)$ is obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}
Regarding the construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation, which we have just revisited.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}

\subsection{Geometrization of characters of the ingredient representations}

\todo{Everything below here is under construction}
Recall that $\sigma^0$ is a representation of $(\underline{G}^0)_\Fq^\reductive$.
Suppose $\pi_0((\underline{G}^0)_\Fq^\reductive)$ is cyclic.
Let $A^0$ be a Weil sheaf on $(\underline{G}^0)_\Fq^\reductive$ such that, as a sheaf on $(\underline{G}^0)_\Fq^\reductive \otimes \bFq$, it is a character sheaf.
The existence of $A^0$ follows from \cite{Lusztig:CSDG}*{}.
Let $A$ be the Weil sheaf on $\underline{G}^0_\Fq$ obtained by pullback along $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.

Let $n$ be an integer greater than the supremum of the depths of $\varphi_i : \underline{G}^i(\OK)\to \CC^\times$, for $i=0,\ldots, d$.
Then, for $i=0,\ldots, d$ we have the commuting diagram 
\[
\begin{tikzcd}
\end{tikzcd}
\]

@@

\begin{lemma}
Let $\underline{G}$ be a smooth group scheme over $\OK$.
Let $\varphi : \underline{G}(\OK) \to \CC^\times$ be a quasicharacter/continuous linear character. 
The there is a Weil sheaf $\cs{F}$ on $\Gr(\underline{G})$ such that $\trFrob{\cs{F}} = \trace \varphi$.
\end{lemma}

\subsection{Geometrization}\label{ssec:geometrization}

The first step in the geometrization of the type $(\,^\circ K,\,^\circ \rho)$ is to replace $\,^\circ K$ with a group scheme $G$ over $\Fq$ so that $G(\Fq) = \,^\circ K$, canonically. 
Set $S = \Spec{\OK}$ and $S_n = \Spec{\OK/\mathfrak{m}^n}$, where $\mathfrak{m}_K$ is the maximal ideal in $\OK$, and $n$ is a positive integer $n$.
Let $G_n$ be the Greenberg transform of $\underline{G} \times_{S} S_n$.
Then $G_n$ is a smooth group scheme over $\Fq$ and 
\[
G_n(\Fq) = \underline{G}(\OK/\mathfrak{m}_K^n),
\]
canonically.
%
Let $G$ be the Greenberg transform of $\underline{G}$.
Then $G$ is a group scheme over $\Fq$ and 
\[
G(\Fq) = \,^\circ K,
\]
canonically.

\begin{proposition}\label{prop:types}
Let $(\,^\circ K,\,^\circ \rho)$ be a Yu type.
With notation as above, suppose $\pi_0(\underline{G}_s\red)$ is cyclic and that $\pi_0((\underline{G}^{j-1})_s\red)$ is cyclic for $j=1,\ldots, d-1$.
Then there is a virtual admissible $\ell$-adic sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\mathcal{F}} = \trace \rho$ under the canonical isomorphism $G(\Fq) = \underline{G}(\OK)$.
\end{proposition}

>>>>

\subsection{Geometrization of quasicharacters of smooth group schemes over $p$-adic fields}

%We begin with the case of continuous linear characters.

Set $S = \Spec{\OK}$ and $S_n = \Spec{\OK/\mathfrak{m}^n}$, where $\mathfrak{m}_K$ is the maximal ideal in $\OK$, and $n$ is a positive integer $n$.
Let $\underline{G}$ be a smooth group scheme over $\OK$.
Let $G_n$ be the Greenberg transform of $\underline{G} \times_{S} S_n$.
Then $G_n$ is a smooth group scheme over $\Fq$ and 
\[
G_n(\Fq) = \underline{G}(\OK/\mathfrak{m}_K^n),
\]
canonically.
%
Let $G$ be the Greenberg transform of $\underline{G}$.
Then $G$ is a group scheme over $\Fq$ and 
\[
G(\Fq) = \underline{G}(\OK)
\]
canonically.

\begin{proposition}\label{prop:quasicharacters}
Let $\underline{G}$ be a smooth group scheme over $\OK$.
Let $\chi : \underline{G}(\OK) \to \CC^\times$ be a quasicharacter/continuous linear character. 
The there is a Weil sheaf $\cs{F}$ on $\Gr(\underline{G})$ such that $\trFrob{\cs{F}} = \trace \chi$.
\end{proposition}

\begin{proof}
\end{proof}

@@

For use in the next section, we recall some basic facts about the Heisenberg-Weil representation.

@@

We may now revisit the ingredients in the construction of the representation $\,^\circ \rho$ of $\underline{G}(\OK)$ along the lines indicated by Yu and recalled at the beginning of this section.
\begin{enumerate}
\item The continuous representation $\varrho$ of $\,^\circ K^0$ is a representation of $\underline{G}^0(\OK)$ obtained by inflation along $\underline{G}^0(\OK) \to \underline{G}^0(\Fq)$. 
Let $\tau$ be that representation of $\underline{G}^0(\Fq) = \underline{G}^0_\Fq(\Fq)$.
In fact, $\tau$ is itself obtained by pulling back a representation $\sigma$ along the $\Fq$-points of the quotient $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
\item Quasicharacters $\varphi_i : \underline{G}^i(\OK) \to \CC^\times$, for $i=0, \ldots d$.
\item The representation $h_{i+1}^*(W_{i+1})$ of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK)$ is obtained by pulling back a representation along 
\[
(\underline{G}^i \ltimes \underline{J}^{i+1})(\OK) \to (\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq).
\]
Let $w_{i+1}$ be that representation of $(\underline{G}^i \ltimes \underline{J}^{i+1})(\Fq) = (\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq)(\Fq)$. 
Then $w_{i+1}$ is itself obtained by pulling back the representation $W_{i+1}$ along the $\Fq$-points of the quotient
\[
\underline{G}_\Fq^i \ltimes \underline{J}^{i+1}_\Fq \to 
\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp.
\]
\end{enumerate}
Regarding the construction of the type $(\,^\circ  K,\,^\circ \rho)$, the paper \cite{yu:models} concludes with the following provocative observation, which we have just revisited.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations are on groups of
 the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.} \cite{yu:03a}
\end{quotation}

In fact, Yu's description of how to assemble $\,^\circ \rho$ from $(\varrho, \varphi_0, \ldots , \varphi_d)$ is slightly different, but equivalent, to what we have done here \cite{Yu:models}*{}.

\subsection{Geometrization of characters of types}\label{ssec:geometrization}

\todo{Everything below here is under construction}
Recall from Section~\ref{ssec:models} that the ingredient representations in the construction of any Yu type are:
\begin{enumerate}
\item A representation $\sigma$ of the $\Fq$-points of the reductive group $(\underline{G}^0)_\Fq^\reductive$ over $\Fq$. 
Then $\tau$ is the representation of the $\Fq$-points of the special fibre of the smooth model $\underline{G}^0$ over $\OK$ and the representation $\varrho$ of $\underline{G}(\OK)$ is produced from $\tau$ by inflation.
\item For $i=0, \ldots, d$, a quasicharacter (continuous linear character) $\varphi_i$ of $\underline{G}^i(\OK)$. 
\item For $i=0, \ldots, d-1$, a representation $W_{i+1}$ of the the Jacobi group $\Sp(V_{i+1}) \ltimes V_{i+1}^\sharp$, viewed as the $\Fq$-points of an algebraic group over $\Fq$.
Then $w_{i+1}$ is the representation of the $\Fq$-points of the special fibre of $\underline{G}^i \ltimes \underline{J}^{i+1}$ over $\OK$ and the representation $h^*(W_{i+1})$ is produced from $w_{i+1}$ by inflation.
\end{enumerate}

We now explain how to geometrize the characters of these ingredient representations.
Suppose $\pi_0((\underline{G}^0)_\Fq^\reductive)$ is cyclic.
Let $A$ be a Weil sheaf on $(\underline{G}^0)_\Fq^\reductive$ such that, as a sheaf ${\bar A}$ on $(\underline{G}^0)_\Fq^\reductive \otimes \bFq$, it is a character sheaf, and
\[
\trFrob{\bar A} = \trace \sigma.
\]
The existence of $A$ follows from \cite{Lusztig:CSDG}*{}.
Let $B$ be the Weil sheaf on $\underline{G}^0_\Fq$ obtained by pullback along $\underline{G}^0_\Fq \to (\underline{G}^0)_\Fq^\reductive$.
Then $\trFrob{\bar B} = \trace \tau$ so
\[
\trace \varrho(k) = \trFrob{\bar B}(k_\Fq)
\]
where $k \mapsto k_\Fq$ refers to the map $\underline{G}^0(\OK) \to \underline{G}^0(\Fq)$.
Observe that the special fibre $\underline{G}^0_\Fq$ of $\underline{G}^0$ is $\Gr_1(\underline{G}^0)$.
Let $\cs{B}$ be the Weil sheaf on the $\Fq$-scheme $\Gr(\underline{G}^0)$ obtained by pulling back $B$ along  $\Gr(\underline{G}^0) \to \Gr_1(\underline{G}^0)$.

For $i=0,\ldots, d-1$, let $\cs{F}_{i+1}$ be the Weil sheaf on the $\Fq$-scheme $\Gr(\underline{G}^i)$ obtained by applying Proposition~\ref{prop:quasicharacters} to $\varphi_i : \underline{G}^i(\OK) \to \CC^\times$.

For $i=0,\ldots, d-1$, let $\cs{T}_i$ be the Weil sheaf on the $\Fq$-scheme $\Gr(\underline{G}^i \ltimes \underline{J}^{i+1})$

@@

\begin{proposition}\label{prop:types}
Let $(\,^\circ K,\,^\circ \rho)$ be a Yu type.
With notation as above, suppose $\pi_0((\underline{G}^0)_\Fq^\reductive)$ is cyclic.
Then there is a virtual admissible Weil sheaf $\mathcal{F}$ on $G$ such that $\trFrob{\mathcal{F}} = \trace \rho$ under the canonical isomorphism $G(\Fq) = \underline{G}(\OK)$.
\end{proposition}
